---
title: "Homework 3"
output: pdf_document
---

## Name: Yucheng Yang

## I worked with: 

**Click the "Knit" button in RStudio to knit this file to a pdf.**

--------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, 
                      warning = FALSE, message = FALSE)
# add packages here 
library(dplyr)
library(ggplot2)
library(babynames)
```




## Problem 1: explain command (no R needed)

### a. 

*answer:*

 m | w 
-- | --
10 | 4
 3 | 1
 7 | 3


### b. 

*answer:* 
classType | m | w | ratioW
------ | -- | -- | --
C | 10 | 4 | 0.16
C | 3 | 1 | 0.04
C | 7 | 3 | 0.12
S | 2 | 7 | 0.28
S | 7 | 10 | 0.4




```{r}
4/25
1/25
3/25
7/25
10/25
```


### c. 

*answer:* 
classType | m | w | ratioW
------ | -- | -- | --
C | 10 | 4 | 0.5
C | 3 | 1 | 0.125
C | 7 | 3 | 0.375
S | 2 | 7 | 0.412
S | 7 | 10 | 0.59
```{r}
4/(4+1+3)
1/(4+1+3)
3/(4+1+3)
7/17
10/17
```


### d. 

*answer:* write your answer here and your Markdown table below

     Y
   | --
 C | 28
 S | 26
 
```{r}
10+4+3+1+7+3
2+7+7+10
```
 


### e. 

*answer:* write your answer here and your Markdown table below
classType | m | w | X | Y
------ | -- | -- | -- | -- 
C | 10 | 4 | 14 | 28
C | 3 | 1 | 4 | 28
C | 7 | 3 | 10 | 28
S | 2 | 7 | 9 | 26
S | 7 | 10 | 17 | 26


------------------------------------------------

## Problem 2: explain error (no R needed)

*answer:* write your answer here 

Variable X=w+m has an unique for every class but Y=sum(w+m) has unique value for every class type, so X and Y have different row number, which is something summarize command don't know how to deal with.

------------------------------------------------

## Problem 3 (give commands, no R needed)

### a. 

*answer:* write your answer below

```
mydata %>% 
  mutate(prop = w/(w+m))
```

### b. 

*answer:* write your answer below

```
mydata %>% 
  mutate(all=sum(m+w))
  group_by(classType) %>%
  mutate(prop = sum(w)/all)
```

------------------------------------------------

## Problem 4: babynames 

### a. 

*answer:* write your answer here
The value of `nrow(babynames_Kat)` is 138, representing that the name Katherine appeared in 138 years of the records between 1880 to 2017. 
```{r}
a<-babynames_Kat<-babynames %>% 
  filter(name=="Katherine", sex=="F") 
nrow(babynames_Kat)
```

### b. 


*answer:* write your answer here
The value is 552, representing that names "Katherine","Catherine", "Katie" or "Kate" owned by females appeared every year in the record from 1880 to 2017. 
```{r}
babynames_Kat<-babynames %>% 
  filter(name%in%c("Katherine","Catherine","Katie","Kate"), sex=="F") 
nrow(babynames_Kat)
```

### c. 

*answer:* write your answer here
The name "Katherine" were used most frequently in 1920,1950 and 1990, and fluctuate in general from 1880 to 2017;
The name "Catherine" were used most frequently in 1920,1958, and fluctuate in general from 1880 to 2017;
The name "Katie" were used most frequently in 1982, before that it weren't used much compared to the above two;
The name "Kate" were used more frequently after 1980, before that it weren't used much compared to the other three.

```{r}
babynames_Kat<-babynames %>% 
  filter(name%in%c("Katherine","Catherine","Katie","Kate"), sex=="F") %>% 
  group_by(year) %>% 
  mutate(tot=sum(n))
Katherine<-babynames %>% 
  filter(name=="Katherine", sex=="F") %>% 
  group_by(year) %>% 
 mutate(tot1=sum(n))
Catherine<-babynames %>% 
  filter(name=="Catherine", sex=="F") %>% 
  group_by(year) %>% 
 mutate(tot2=sum(n))
Katie<-babynames %>% 
  filter(name=="Katie", sex=="F") %>% 
  group_by(year) %>% 
 mutate(tot3=sum(n))
Kate<-babynames %>% 
  filter(name=="Kate", sex=="F") %>% 
  group_by(year) %>% 
 mutate(tot4=sum(n))
  ggplot()+geom_point(Katherine, mapping= aes(x=year,y=tot1),color="blue")+geom_point(Catherine, mapping= aes(x=year,y=tot2),color="red")+geom_point(Katie, mapping= aes(x=year,y=tot3),color="green")+geom_point(Kate, mapping= aes(x=year,y=tot4))
  scale_x_continuous(breaks=seq(1880,2017,by=10))
```

### d. 


*answer:* write your answer here
Blair and Elisa have names with female/male proportions closest to 0.5.
```{r}
babynames %>% 
  group_by(sex,name) %>% 
  summarise(eachtot=sum(n)) %>% 
  filter(eachtot >= 10000) %>% 
  group_by(name) %>% 
  mutate(bal=eachtot/sum(eachtot)) %>% filter(bal>0.495,bal<0.505)
  
```


------------------------------------------------

## Problem 5: restaurant violations

### a. 

*answer:* 


```{r}
data("Violations", package = "mdsr")
p5<-Violations%>%
  tidyr::drop_na(score)%>%
  filter(boro == "MANHATTAN")%>%
  dplyr::group_by(zipcode)%>%
  dplyr::summarize(inspections = n(), mscore = median(score)) %>%
  filter(inspections >= 50)
p5
  
```

### b. 

*answer:* write your answer here
It seems that when there are less inspections done, the median score tend to have more variability; when there's more inspections done, the median score tend to to more stable and mostly fluctuate between 15 to 18.
```{r}
ggplot(p5,aes(x=inspections,y=mscore))+geom_point()+geom_smooth(se=FALSE)+ggtitle("Restaurant Violations in Manhattan")+xlab("Number of inspections")+ylab("Median score")
```


------------------------------------------------


## Problem 6: flights

Use the `flights` data frame from the `nycflights13` package to answer the questions below. (see help `?nycflights13::flights` for more details) Use `dplyr` to answer the questions below. 
### a.
*answer:* write your answer here

Plane N725MQ traveled the most from NYC airports in 2013.
```{r}
data("flights", package = "nycflights13")
flights %>% group_by(tailnum) %>% 
  summarise(ct=n()) %>% slice_max(ct,n=5)
```

### b. 
*answer:* write your answer here
We tend to see most flights from NYC on July since July has the highest median number of flights per day.
```{r}
library(lubridate)
s1<-flights %>% mutate(mth = month(time_hour, label = TRUE)) %>% group_by(month,day) %>% mutate(cts=n())
s1
ggplot(s1,aes(x=mth,y=cts))+geom_boxplot()
```

### c. 

*answer:* write your answer here
Saturdays tend to have the fewest number of flights from NYC.
```{r}
flights %>% mutate(dow = wday(time_hour, label = TRUE)) %>% 
  group_by(month,day) %>% 
  mutate(cts2=n()) %>% 
  ggplot(aes(x=dow,y=cts2))+geom_boxplot()
```
