---
title: "Homework 6"
output: pdf_document
---

## Name: Yucheng Yang

## I worked with: 


--------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, 
                      warning = FALSE, message = FALSE)
library(tidyverse) 
library(tidyr)
library(nasaweather)
library(purrr)
library(alphavantager)
library(ggplot2)
library(rvest)
library(stringr)
```

## Assignment prompt

## Problem 1: IPS

### a.


*answer:*

```{r}
address<-readr::read_lines("https://math.carleton.edu/kstclair/data/mac-address.txt")
```

### b.


*answer:*

```{r}
main<-address %>% str_split(";") %>% unlist()
main
```

### c.

*answer:*

```{r}
mt<-str_split(main[1:4], "=", simplify = TRUE)
names<-mt[1:4, 1]
values<-mt[1:4, 2]
names
values
```

### d.

*answer:*

```{r}
rest<-str_split(main[5:15], "=", simplify = TRUE) 
a<-rest[1:11,1]
b<-rest[1:11,2] %>% str_split( ",", simplify = TRUE)
info<-as.matrix(cbind(a,b))
info
```

### e.

*answer:*

```{r}
colnames(info)<-c("mac", "signal", "channel", "type")
info_n<-as.data.frame(info)
offline <- info_n %>% mutate(t=values[1], id=values[2], pos=values[3], degree=values[4])
offline<-offline[,c(5,6,7,8,1,2,3,4)] %>% as_tibble()
offline
```

### f

*answer:*

```{r}
offline <- offline %>% separate(pos, c("pos_x", "pos_y", "pos_z"), ",", convert=TRUE) 
offline$degree<-as.numeric(offline$degree)
offline$signal<-as.numeric(offline$signal)
offline
```


---------------------------------------

## Problem 2: weather
Load the `nasaweather` data and look at the help file for the atmospheric data `?atmos`. 

### a.

*answer:*

```{r}
library(nasaweather)
nasaweather::atmos
coef<-rep(NA,7)
coef_vari<-function(x) {
  for (i in 1:7){
    coef[i]<-sd(x[[i]], na.rm = TRUE)/mean(x[[i]], na.rm = TRUE)
  }
coef
}
coef_vari(atmos[5:11])
```

### b.Use a `map` function (from `purrr` package) to compute the coefficient of variation for all atmospheric measurements except location (lat/long) and time (year, month) variables. Use a function that returns a vector or data frame (and show this output).

*answer:*

```{r}
atmos[,5:11] %>% 
  map_df(function(x){
    tibble(
      coef_var=sd(x,na.rm=TRUE)/mean(x,na.rm=TRUE)
    )
  })
```

### c.Create a function called `my_stats` that computes the following statistics for an input vector: mean, sd, and 5-number summary (min/Q1/median/Q3/max). The function should return an  output object that is a "named" vector. E.g. the following is a named vector with the name of the left side of `=` and the value on the right side. Show the output of your function by inputting the `temp` values from `atmos`.

*answer:*

```{r}
#c(x = 1, y = 2)
my_stats <- function(x) {
  mn<-mean(x,na.rm=TRUE)
  standard_dev<-sd(x,na.rm=TRUE)
  minimum<-min(x,na.rm=TRUE)
  q1<-quantile(x,0.25,na.rm=TRUE)
  med<-median(x,na.rm=TRUE)
  q3<-quantile(x,0.75,na.rm=TRUE)
  maximum<-max(x,na.rm=TRUE)
  c(mean=mn,sd=standard_dev,min=minimum,Q1=q1,median=med,Q3=q3,max=maximum)
  
}
my_stats(atmos$temp)
```

### d.


*answer:*

```{r}
atmos[,5:11] %>% 
  map_df(my_stats) %>% mutate(var_name = colnames(atmos[,5:11]))
```

### e.

*answer:*

```{r}
atmos %>% 
  group_by(year) %>% 
  summarise( try = my_stats(temp)) %>% 
  mutate(var_name=c("mean", "sd", "min", "Q1", "median", "Q3", "max")) %>% 
  pivot_wider(names_from = year, values_from = try)
```


---------------------------------------

## Problem 3: Crimes



```{r, eval = FALSE}
ggplot(crimes, aes(x = violent_crime, y = property_crime, label = City)) +
    geom_point() +
    geom_text(
      data = filter(crimes_clean, violent_crime > 1500 | property_crime > 6500), 
      check_overlap = TRUE, size = 2.5, nudge_y = 40
    )
```



*answer:*

```{r}
#sorry my network always fails when retrieving html file from the internet, so i had to download the html page and put it in the directory for the code to run
url<- "crime_rate_wiki.htm"
page <- read_html(url)
tables.html <- html_nodes(page, "table")
#str(tables.html)
table<-html_table(tables.html[[1]], fill=TRUE)
crime<- table %>% select(c(2,4,10)) 
colnames(crime)[1]<-"City"
colnames(crime)[2]<-"violent_crime"
colnames(crime)[3]<-"property_crime"
crime
new_crime<-crime[-1, ] %>% 
  drop_na() %>%
  mutate(violent_crime=parse_number(violent_crime), property_crime=parse_number(property_crime))
new_crime <- new_crime %>% as.tibble() %>% drop_na()
ggplot(new_crime, aes(x = violent_crime, y = property_crime, label = City)) +
    geom_point() +
    geom_text(
      data = filter(new_crime, violent_crime > 1500 | property_crime > 6500), 
      check_overlap = TRUE, size = 2.5, nudge_y = 40
    )
```


----------------------------------------

## Problem 4: Movie scraping


### a.

*answer:*
There is one HTML table on the page that contains the box office earnings. 
```{r}
url1 <- "https://www.boxofficemojo.com/chart/ww_top_lifetime_gross/?offset=0&area=XWW"
page1 <- read_html(url1)
str(page1)
tables.html1 <- html_nodes(page1, "table")
str(tables.html1)
html_text(tables.html1)
```



### b.

*answer:*

```{r}
tables<-html_table(tables.html1[[1]], fill=TRUE)
glimpse(tables)
movie<-tables %>% as.tibble()
colnames(movie)[1:8]<-c("rank", "title", "world_dollars","domestic_dollars", "domestic_percentage", "overseas_dollars", "overseas_percentage", "year")
movie
```

### c.

*answer:*

```{r}
movie<-movie %>% 
  mutate(world_dollars=parse_number(world_dollars), domestic_dollars = parse_number(domestic_dollars), domestic_percentage = parse_number(domestic_percentage), overseas_dollars = parse_number(overseas_dollars), overseas_percentage= parse_number( overseas_percentage))
movie
```

### d.


*answer:*

```{r}
links_table<-html_nodes(tables.html1[[1]], "a")
pulled<-html_attr(links_table[5], "href")
full<-str_c("http://www.boxofficemojo.com", pulled)
full
```

### e.

*answer:*


```{r}
first_page <- "https://www.boxofficemojo.com/chart/ww_top_lifetime_gross/?offset=800&area=XWW"
read_page <- function(x){
  page <- read_html(x)
  tables.html <- html_nodes(page, "table")
  table<- html_table(tables.html1[[1]], fill=TRUE) %>% as.tibble()
  colnames(table)[1:8]<-c("rank", "title", "world_dollars","domestic_dollars", "domestic_percentage", "overseas_dollars", "overseas_percentage", "year")
  if(is.character(table$rank)){
    table %>% mutate(rank=as.character(rank)) %>% 
      mutate(world_dollars=parse_number(world_dollars), domestic_dollars = parse_number(domestic_dollars), domestic_percentage = parse_number(domestic_percentage), overseas_dollars = parse_number(overseas_dollars), overseas_percentage= parse_number( overseas_percentage))
  }
  else{
    table %>% mutate(world_dollars=parse_number(world_dollars), domestic_dollars = parse_number(domestic_dollars), domestic_percentage = parse_number(domestic_percentage), overseas_dollars = parse_number(overseas_dollars), overseas_percentage= parse_number( overseas_percentage))
  }
  table
}
read_page(first_page)
```


### f.

*answer:*

```{r}
temp_url <- "https://www.boxofficemojo.com/chart/ww_top_lifetime_gross/?offset=#&area=XWW"
five<-rep(temp_url,5)
five[1]<-str_replace(five[[1]], "#", "0")
five[2]<-str_replace(five[[2]], "#", "200")
five[3]<-str_replace(five[[3]], "#", "400")
five[4]<-str_replace(five[[4]], "#", "600")
five[5]<-str_replace(five[[5]], "#", "800")
five
five_page<-map_df(five, read_page)
glimpse(five_page)
```
