---
title: "Error Term Simulation"
author: "Muyang Shi"
date: "10/22/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading-data-and-library}
library(lme4)
library(HLMdiag)
library(MASS)
library(dplyr)
radon <- HLMdiag::radon
library(lmtest)
```

### Utility
```{r bptest}
bp <- function (resid,fitted){
    .sq_resid <- resid^2 # squared residual
    
    # Regression e^2 on the fitted values
    .mod <-lm(.sq_resid ~ fitted)
    
    # Regression sum of squares
    .SSR <- sum((predict(.mod) - mean(.sq_resid))^2) # anova(.mod)[1,2]
    # Error Sum of Squares
    .SSE <- sum(.sq_resid) # anova(.mod)[2,2]
    # n
    .n <- length(resid)
    # test statistics
    .x <- (.SSR/2) / (.SSE/.n)^2
    
    .p <- pchisq(.x,df=1)
    .p <- ifelse(.p>=0.5, 1-.p,.p)
    return(c(stats=.x,p_val=.p))
}
```

```{r brown-forsythe-test}
bf <- function (resid, fitted){
    .frame <- data.frame(resid=resid,fitted=fitted)
    # .frame <- .frame[order(.frame$fitted),]
    .frame$group <- ifelse(.frame$fitted <= median(.frame$fitted),1,2)
    # Split the data into two pieces
    .group1 <- .frame[c(which(.frame$group==1)),]
    .group2 <- .frame[c(which(.frame$group==2)),]
        
    # Calculate the medians of each group
    .med1 <- median(.group1$resid)
    .med2 <- median(.group2$resid)
    
    # Calculate the absolute deviations of the residuals from the median residual within each group
    .AD1 <- abs(.group1$resid - .med1)
    .AD2 <- abs(.group2$resid - .med1)
    ttest <- t.test(.AD1,.AD2,var.equal=TRUE)
    return(c(stats=ttest$statistic,p_val=ttest$p.value))
}
```

```{r brown-forsythe-test-4-bins}
bf4 <- function (resid, fitted){
    .frame <- data.frame(resid=resid,fitted=fitted)
    .quantiles <- quantile(.frame$fitted)
    .q1 <- .quantiles[2]
    .med <- .quantiles[3]
    .q3 <- .quantiles[4]
    .frame$group <- ifelse(.frame$fitted <= .q1, 1,
                        ifelse(.frame$fitted <= .med, 2,
                            ifelse(.frame$fitted <= .q3, 3, 4)))

    # Split the data into four pieces
    .group1 <- .frame[c(which(.frame$group==1)),]
    .group2 <- .frame[c(which(.frame$group==2)),]
    .group3 <- .frame[c(which(.frame$group==3)),]
    .group4 <- .frame[c(which(.frame$group==4)),]

    .med1 <- median(.group1$resid)
    .med2 <- median(.group2$resid)
    .med3 <- median(.group3$resid)
    .med4 <- median(.group4$resid)

    # absolute deviance
    .AD1 <- abs(.group1$resid - .med1)
    .AD2 <- abs(.group2$resid - .med2)
    .AD3 <- abs(.group3$resid - .med3)
    .AD4 <- abs(.group4$resid - .med4)

    # Choose the most extreme difference between avg of AD

    .mean1 <- sum(.AD1)/length(.AD1)
    .mean2 <- sum(.AD2)/length(.AD2)
    .mean3 <- sum(.AD3)/length(.AD3)
    .mean4 <- sum(.AD4)/length(.AD4)

    .minADindex <- which.min(c(.mean1,.mean2,.mean3,.mean4))
    .maxADindex <- which.max(c(.mean1,.mean2,.mean3,.mean4))

    .minAD <- list(.AD1,.AD2,.AD3,.AD4)[[.minADindex]]
    .maxAD <- list(.AD1,.AD2,.AD3,.AD4)[[.maxADindex]]

    ttest <- t.test(.minAD,.maxAD,var.equal=TRUE)
    return(c(stats=ttest$statistic,p_val=ttest$p.value))
}
```

### Simulation Function(s)

Simulation function that assume multivariate normal covariance between $b_0$ and $b_1$, the random intercept and random slope.

#### Error term normality assumption
```{r simulation-function-normality}
sim.hlm <- function(.mod, nsim, e.dsn) {
    vc <- VarCorr( .mod )
    D  <- as.matrix( bdiag(vc) )
    
    sigma.err <- attr(vc, "sc") # std of the error term
    
    n <- getME(.mod,"n") # number of error terms (number of observations)
    m <- getME(.mod,"q") / nrow(D) # number of random slope or random intercepts
    
    ## Simulating error terms
    if(e.dsn == "norm") {
    	e  <- rnorm(n = nsim * n, mean = 0, sd = sigma.err)
    } 
    if(e.dsn == "t") {
    	e  <- (sigma.err / sqrt(3)) * rt(n = nsim * n, df = 3)
    }
    if(e.dsn == "exp") {
    	e  <- sigma.err * ( rexp(n = nsim * n) - 1 )
    }
    e <- matrix(e, nc = nsim)
    
    ## Simulating random intercept
    b.sim <- mvrnorm(n = nsim * m, mu = c(0, 0), Sigma = D) # allowing for covariance between b0 and b1
    
    b0 <- matrix(b.sim[,1], nc = nsim)
    b1 <- matrix(b.sim[,2], nc = nsim)
    
    ## Generating y
    b <- rbind(b0, b1)
    y <- getME(.mod, "X") %*% fixef(.mod) + getME(.mod, "Z") %*% b + e
    
    y.df <- as.data.frame( as.matrix( y) )
    colnames(y.df) <- paste("sim_", 1:ncol(y.df), sep = "")
    
    return( y.df )
}
```

#### Error Term Constant Variance Assumption
```{r simulation-function-constant-variance}
sim.hlmvar <- function(.mod, nsim, evar.dsn) {
    vc <- VarCorr( .mod )
    D  <- as.matrix( bdiag(vc) )
    
    sigma.err <- attr(vc, "sc") # std of the error term
    
    n <- getME(.mod,"n") # number of error terms (number of observations)
    m <- getME(.mod,"q") / nrow(D) # number of random slope or random intercepts
    
    ## Simulating random intercept and slope
    b.sim <- mvrnorm(n = nsim * m, mu = c(0, 0), Sigma = D) # allowing for covariance between b0 and b1
    
    b0 <- matrix(b.sim[,1], nc = nsim)
    b1 <- matrix(b.sim[,2], nc = nsim)
    
    ## Generating y's
    b <- rbind(b0, b1)
    y <- getME(.mod, "X") %*% fixef(.mod) + getME(.mod, "Z") %*% b # True y without error term
    y_backup <- y # Save the true y value

    ## Simulating error terms
    ## And append the simulated error terms to get the simulated y values
    if(evar.dsn == "-") { # constant variance
    	e  <- rnorm(n = nsim * n, mean = 0, sd = sigma.err)
    	y <- y + e
    } 
    if(evar.dsn == ">") { # big variance on the left, and small variance on the right
    	for(i in 1:ncol(y)){
    	    # Preparation work, split true y's into four bins
    	    y_col <- sort(y[,i]) # Sort the true y's
    	    total_num <- length(y_col)
    	    bin_width <- round(total_num/4)
    	    bin_1 <- y_col[c(1:bin_width)]
    	    bin_2 <- y_col[c((bin_width+1):(2*bin_width))]
    	    bin_3 <- y_col[c((2*bin_width+1):(3*bin_width))]
    	    bin_4 <- y_col[c((3*bin_width+1):total_num)]
    	    
    	    # Simulating non-constant variance error terms
    	    e_1 <- rnorm(n = bin_width, mean = 0, sd = sqrt(7*sigma.err^2/16))
    	    e_2 <- rnorm(n = bin_width, mean = 0, sd = sqrt(5*sigma.err^2/16))
    	    e_3 <- rnorm(n = bin_width, mean = 0, sd = sqrt(3*sigma.err^2/16))
    	    e_4 <- rnorm(n = length(bin_4), mean = 0, sd = sqrt(1*sigma.err^2/16))
    	    
    	    # Adding simulated error term to the true value
    	    ysim_bin_1 <- bin_1 + e_1
    	    ysim_bin_2 <- bin_2 + e_2
    	    ysim_bin_3 <- bin_3 + e_3
    	    ysim_bin_4 <- bin_4 + e_4
    	    # Join the binned simulated y's back together
    	    y_simulated_col <- c(ysim_bin_1,ysim_bin_2,ysim_bin_3,ysim_bin_4)
    	    # Update the simulation matrix (column)
    	    y[,i] <- y_simulated_col
    	}
    }
    if(evar.dsn == "<") { # small variance on the left, and big variance on the right
    	for(i in 1:ncol(y)){
    	    # Preparation work, split true y's into four bins
    	    y_col <- sort(y[,i]) # Sort the true y's
    	    total_num <- length(y_col)
    	    bin_width <- round(total_num/4)
    	    bin_1 <- y_col[c(1:bin_width)]
    	    bin_2 <- y_col[c((bin_width+1):(2*bin_width))]
    	    bin_3 <- y_col[c((2*bin_width+1):(3*bin_width))]
    	    bin_4 <- y_col[c((3*bin_width+1):total_num)]
    	    
    	    # Simulating non-constant variance error terms
    	    e_1 <- rnorm(n = bin_width, mean = 0, sd = sqrt(1*sigma.err^2/16))
    	    e_2 <- rnorm(n = bin_width, mean = 0, sd = sqrt(3*sigma.err^2/16))
    	    e_3 <- rnorm(n = bin_width, mean = 0, sd = sqrt(5*sigma.err^2/16))
    	    e_4 <- rnorm(n = length(bin_4), mean = 0, sd = sqrt(7*sigma.err^2/16))
    	    
    	    # Adding simulated error term to the true value
    	    ysim_bin_1 <- bin_1 + e_1
    	    ysim_bin_2 <- bin_2 + e_2
    	    ysim_bin_3 <- bin_3 + e_3
    	    ysim_bin_4 <- bin_4 + e_4
    	    # Join the binned simulated y's back together
    	    y_simulated_col <- c(ysim_bin_1,ysim_bin_2,ysim_bin_3,ysim_bin_4)
    	    # Update the simulation matrix (column)
    	    y[,i] <- y_simulated_col
    	}
    }
    if(evar.dsn == "<>"){ # small variance on the end, big variance in the middle
    	for(i in 1:ncol(y)){
    	    # Preparation work, split true y's into four bins
    	    y_col <- sort(y[,i]) # Sort the true y's
    	    total_num <- length(y_col)
    	    bin_width <- round(total_num/4)
    	    bin_1 <- y_col[c(1:bin_width)]
    	    bin_2 <- y_col[c((bin_width+1):(2*bin_width))]
    	    bin_3 <- y_col[c((2*bin_width+1):(3*bin_width))]
    	    bin_4 <- y_col[c((3*bin_width+1):total_num)]
    	    
    	    # Simulating non-constant variance error terms
    	    e_1 <- rnorm(n = bin_width, mean = 0, sd = sqrt(2*sigma.err^2/16))
    	    e_2 <- rnorm(n = bin_width, mean = 0, sd = sqrt(6*sigma.err^2/16))
    	    e_3 <- rnorm(n = bin_width, mean = 0, sd = sqrt(6*sigma.err^2/16))
    	    e_4 <- rnorm(n = length(bin_4), mean = 0, sd = sqrt(2*sigma.err^2/16))
    	    
    	    # Adding simulated error term to the true value
    	    ysim_bin_1 <- bin_1 + e_1
    	    ysim_bin_2 <- bin_2 + e_2
    	    ysim_bin_3 <- bin_3 + e_3
    	    ysim_bin_4 <- bin_4 + e_4
    	    # Join the binned simulated y's back together
    	    y_simulated_col <- c(ysim_bin_1,ysim_bin_2,ysim_bin_3,ysim_bin_4)
    	    # Update the simulation matrix (column)
    	    y[,i] <- y_simulated_col
    	}
    }
    if(evar.dsn == "><"){
    	for(i in 1:ncol(y)){
    	    # Preparation work, split true y's into four bins
    	    y_col <- sort(y[,i]) # Sort the true y's
    	    total_num <- length(y_col)
    	    bin_width <- round(total_num/4)
    	    bin_1 <- y_col[c(1:bin_width)]
    	    bin_2 <- y_col[c((bin_width+1):(2*bin_width))]
    	    bin_3 <- y_col[c((2*bin_width+1):(3*bin_width))]
    	    bin_4 <- y_col[c((3*bin_width+1):total_num)]
    	    
    	    # Simulating non-constant variance error terms
    	    e_1 <- rnorm(n = bin_width, mean = 0, sd = sqrt(6*sigma.err^2/16))
    	    e_2 <- rnorm(n = bin_width, mean = 0, sd = sqrt(2*sigma.err^2/16))
    	    e_3 <- rnorm(n = bin_width, mean = 0, sd = sqrt(2*sigma.err^2/16))
    	    e_4 <- rnorm(n = length(bin_4), mean = 0, sd = sqrt(6*sigma.err^2/16))
    	    
    	    # Adding simulated error term to the true value
    	    ysim_bin_1 <- bin_1 + e_1
    	    ysim_bin_2 <- bin_2 + e_2
    	    ysim_bin_3 <- bin_3 + e_3
    	    ysim_bin_4 <- bin_4 + e_4
    	    # Join the binned simulated y's back together
    	    y_simulated_col <- c(ysim_bin_1,ysim_bin_2,ysim_bin_3,ysim_bin_4)
    	    # Update the simulation matrix (column)
    	    y[,i] <- y_simulated_col
    	}
    }
    
    y.df <- as.data.frame( as.matrix( y) )
    colnames(y.df) <- paste("sim_", 1:ncol(y.df), sep = "")
    
    return( y.df )
}
```

#### Error Term Linearity Assumption
```{r simulation-function-linearity}
sim.hlmlin <- function(.mod, nsim, rel.dsn) {
    vc <- VarCorr( .mod )
    D  <- as.matrix( bdiag(vc) )
    
    sigma.err <- attr(vc, "sc") # std of the error term
    
    n <- getME(.mod,"n") # number of error terms (number of observations)
    m <- getME(.mod,"q") / nrow(D) # number of random slope or random intercepts
    
    ## Simulating error terms
    e <- rnorm(n = nsim*n, mean = 0, sd = sigma.err)
    e <- matrix(e, nc = nsim)

    ## Simulating random intercept
    b.sim <- mvrnorm(n = nsim * m, mu = c(0, 0), Sigma = D) # allowing for covariance between b0 and b1
    
    b0 <- matrix(b.sim[,1], nc = nsim)
    b1 <- matrix(b.sim[,2], nc = nsim)
    
    ## Modifying linearity relationship
    X <- getME(.mod, "X")
    if(rel.dsn == "linear") {
        # Do nothing
    } 
    if(rel.dsn == "sqrt") {
        signs <- sign(X[,"uranium"])
        values <- abs(X[,"uranium"])
        X[,"uranium"] <- signs*sqrt(values)
    }
    if(rel.dsn == "sq") {
        signs <- sign(X[,"uranium"])
        values <- abs(X[,"uranium"])
        X[,"uranium"] <- signs*(values^2)
    }
    if(rel.dsn == "log"){
        signs <- sign(X[,"uranium"])
        values <- abs(X[,"uranium"])
        X[,"uranium"] <- signs*(log(values))
    }
    
    ## Generating y
    b <- rbind(b0, b1)
    y <- X %*% fixef(.mod) + getME(.mod, "Z") %*% b + e
    
    y.df <- as.data.frame( as.matrix( y) )
    colnames(y.df) <- paste("sim_", 1:ncol(y.df), sep = "")
    
    return( y.df )
}
```

### Simulations

```{r simulation-setup}
## fitted model
(fm <- lmer(log.radon ~ basement + uranium + (basement | county), data = radon))

# (fm2 <- lmer(log.radon ~ basement + uranium + (1 | county) + (0 + basement | county), data = radon))

## Distributional settings
e.dsn <- c("norm", "t", "exp") # Normality Assumption

evar.dsn <- c("-",">","<","<>","><") # Constant Variance Assumption

rel.dsn <- c("linear","sqrt","sq","log") # Linearity Assumption
```

#### Simulation on non-normality
```{r simulaion-normality}
## Set seed and warning handler
set.seed(7231985)
options(warn = 2) # turns warnings into errors

## Simulating
for(i in seq_along(e.dsn)){
	# Simulate
	sim.fm <- sim.hlm( fm, nsim = 5000, e.dsn = e.dsn[i])
                   	   
	# Excluding those simulations with convergence issues 
	# and replacing them with new simulations

	repeat{
  		sim.models <- lapply(sim.fm, function(x) try(refit(object = fm, newresp = x)))
  
  		conv.issues <- which(sapply(sim.models, function(x) class(x) == "try-error"))
  
  		if(length(conv.issues) == 0) break
  
  		# remove simulations with convergence issues
  		sim.fm <- sim.fm[,-conv.issues]
  
  		if(length(sim.fm) == 5000) break
  
  		# Additional simulations to achieve correct number of simulations
  		sim.fm <- cbind(sim.fm, sim.hlm(fm, nsim = length(conv.issues), e.dsn = e.dsn[i]))
	}
	saveRDS(sim.models, paste("simulated_models/", e.dsn[i], "_models.RDS", sep = ""))
}
```

#### Simulation on non-constant variance
```{r simulation-constant-variance}
## Set seed and warning handler
set.seed(7231985)
options(warn = 2) # turns warnings into errors

## Simulating
for(i in seq_along(evar.dsn)){
	# Simulate
	sim.fm <- sim.hlmvar( fm, nsim = 5000, evar.dsn = evar.dsn[i])
                   	   
	# Excluding those simulations with convergence issues 
	# and replacing them with new simulations

	repeat{
  		sim.models <- lapply(sim.fm, function(x) try(refit(object = fm, newresp = x)))
  
  		conv.issues <- which(sapply(sim.models, function(x) class(x) == "try-error"))
  
  		if(length(conv.issues) == 0) break
  
  		# remove simulations with convergence issues
  		sim.fm <- sim.fm[,-conv.issues]
  
  		if(length(sim.fm) == 5000) break
  
  		# Additional simulations to achieve correct number of simulations
  		sim.fm <- cbind(sim.fm, sim.hlmvar(fm, nsim = length(conv.issues), evar.dsn = evar.dsn[i]))
	}
	saveRDS(sim.models, paste("simulated_models/", evar.dsn[i], "_models.RDS", sep = ""))
}
```

#### Simulation on non-linearity
```{r simulation-linearity}
## Set seed and warning handler
set.seed(7231985)
options(warn = 2) # turns warnings into errors

## Simulating
for(i in seq_along(rel.dsn)){
	# Simulate
	sim.fm <- sim.hlmlin( fm, nsim = 5000, rel.dsn = rel.dsn[i])
                   	   
	# Excluding those simulations with convergence issues 
	# and replacing them with new simulations

	repeat{
  		sim.models <- lapply(sim.fm, function(x) try(refit(object = fm, newresp = x)))
  
  		conv.issues <- which(sapply(sim.models, function(x) class(x) == "try-error"))
  
  		if(length(conv.issues) == 0) break
  
  		# remove simulations with convergence issues
  		sim.fm <- sim.fm[,-conv.issues]
  
  		if(length(sim.fm) == 5000) break
  
  		# Additional simulations to achieve correct number of simulations
  		sim.fm <- cbind(sim.fm, sim.hlmvar(fm, nsim = length(conv.issues), rel.dsn = rel.dsn[i]))
	}
	saveRDS(sim.models, paste("simulated_models/", rel.dsn[i], "_models.RDS", sep = ""))
}
```
### Analysis

#### Normality Assumption
```{r some-scratch-work}
# exp1 <- `exp_model`[[1]]
# plot(exp1)
# # resid(exp1)
# shapiro.test(resid(exp1))
# 
# t1 <- `t_model`[[1]]
# plot(t1)
# # resid(exp1)
# shapiro.test(resid(t1))
# qqnorm(resid(t1))
```

```{r shapiro-tests}
`norm_model` <- readRDS("simulated_models/norm_models.RDS")
`t_model` <- readRDS("simulated_models/t_models.RDS")
`exp_model` <- readRDS("simulated_models/exp_models.RDS")

normal_tests <- lapply(`norm_model`,function(x){
    model_normality_test <- shapiro.test(resid(x))
    # print(model_normality_test)
    })
saveRDS(normal_tests,file="./Result/Normality/normal.RDS")

t_tests <- lapply(`t_model`,function(x){
    model_normality_test <- shapiro.test(resid(x))
    # print(model_normality_test)
    })
saveRDS(t_tests,file="./Result/Normality/t.RDS")

exp_tests <- lapply(`t_model`,function(x){
    model_normality_test <- shapiro.test(resid(x))
    # print(model_normality_test)
    })
saveRDS(exp_tests,file="./Result/Normality/exp.RDS")
```

#### Constant-Variance Assumption
```{r load-heteroscadasticity-data}
`var-` <- readRDS("simulated_models/-_models.RDS")
`var<` <- readRDS("simulated_models/<_models.RDS")
`var>` <- readRDS("simulated_models/>_models.RDS")
`var<>` <- readRDS("simulated_models/<>_models.RDS")
`var><` <- readRDS("simulated_models/><_models.RDS")
```

```{r Breusch-Pagan-Test}
# library(lmtest)
# sim1 <- (`var-`[[1]])
# plot(sim1)
# resid(sim1)
# fitted(sim1)
# plot(fitted(sim1),resid(sim1))
# mod <- lm(resid(sim1) ~ fitted(sim1))
# # plot(mod,1)
# bptest(resid(sim1) ~ fitted(sim1))

`var-tests` <- lapply(`var-`,function(x){
    bp <- bp(resid(x),fitted(x))
    return(bp)
    })
saveRDS(`var-tests`,file="./Result/ConstantVariance/BP/var-.RDS")

`var<tests` <- lapply(`var<`,function(x){
    bp <- bp(resid(x),fitted(x))
    return(bp)
    })
saveRDS(`var<tests`,file="./Result/ConstantVariance/BP/var<.RDS")

`var>tests` <- lapply(`var>`,function(x){
    bp <- bp(resid(x),fitted(x))
    return(bp)
    })
saveRDS(`var>tests`,file="./Result/ConstantVariance/BP/var>.RDS")

`var<>tests` <- lapply(`var<>`,function(x){
    bp <- bp(resid(x),fitted(x))
    return(bp)
    })
saveRDS(`var<>tests`,file="./Result/ConstantVariance/BP/var<>.RDS")

`var><tests` <- lapply(`var><`,function(x){
    bp <- bp(resid(x),fitted(x))
    return(bp)
    })
saveRDS(`var><tests`,file="./Result/ConstantVariance/BP/var><.RDS")
```

```{r Brown-Forsythe-test}
`bf-tests` <- lapply(`var-`,function(x){
    bftest <- bf(resid(x),fitted(x))
    return(bftest)
    })
saveRDS(`bf-tests`,file="./Result/ConstantVariance/BF/var-.RDS")

`bf<tests` <- lapply(`var<`,function(x){
    bftest <- bf(resid(x),fitted(x))
    return(bftest)
    })
saveRDS(`bf<tests`,file="./Result/ConstantVariance/BF/var<.RDS")

`bf>tests` <- lapply(`var>`,function(x){
    bftest <- bf(resid(x),fitted(x))
    return(bftest)
    })
saveRDS(`bf>tests`,file="./Result/ConstantVariance/BF/var>.RDS")

`bf<>tests` <- lapply(`var<>`,function(x){
    bftest <- bf(resid(x),fitted(x))
    return(bftest)
    })
saveRDS(`bf<>tests`,file="./Result/ConstantVariance/BF/var<>.RDS")

`bf><tests` <- lapply(`var><`,function(x){
    bftest <- bf(resid(x),fitted(x))
    return(bftest)
    })
saveRDS(`bf><tests`,file="./Result/ConstantVariance/BF/var><.RDS")
```

```{r Brown-Forsythe4-test}
`bf4-tests` <- lapply(`var-`,function(x){
    bf4test <- bf4(resid(x),fitted(x))
    return(bf4test)
    })
saveRDS(`bf4-tests`,file="./Result/ConstantVariance/BF4/var-.RDS")

`bf4<tests` <- lapply(`var<`,function(x){
    bf4test <- bf4(resid(x),fitted(x))
    return(bf4test)
    })
saveRDS(`bf4<tests`,file="./Result/ConstantVariance/BF4/var<.RDS")

`bf4>tests` <- lapply(`var>`,function(x){
    bf4test <- bf4(resid(x),fitted(x))
    return(bf4test)
    })
saveRDS(`bf4>tests`,file="./Result/ConstantVariance/BF4/var>.RDS")

`bf4<>tests` <- lapply(`var<>`,function(x){
    bf4test <- bf4(resid(x),fitted(x))
    return(bf4test)
    })
saveRDS(`bf4<>tests`,file="./Result/ConstantVariance/BF4/var<>.RDS")

`bf4><tests` <- lapply(`var><`,function(x){
    bf4test <- bf4(resid(x),fitted(x))
    return(bf4test)
    })
saveRDS(`bf4><tests`,file="./Result/ConstantVariance/BF4/var><.RDS")
```

#### Linearity Assumption

```{r load-data}
# normality_norm <- readRDS("norm_models.RDS")
`linear_term_models` <- readRDS("simulated_models/linear_models.RDS")
`sq_term_models` <- readRDS("simulated_models/sq_models.RDS")
```

```{r linearity-test-using-anova}
linearity_linear <- lapply(`linear_term_models`,function(mod){
    model_linear_term <- lmer(log.radon ~ basement + uranium + (basement | county), data = mod@frame)
    model_sq_term <- lmer(log.radon ~ basement + uranium + I(uranium^2) + (basement | county), data = mod@frame)
    anova_result <- anova(model_linear_term,model_sq_term)
    return(c(stats=anova_result$Chisq[2],p_val=anova_result$`Pr(>Chisq)`[2])) # return test stats and p-value
    })
# Model summary can be later checked using summary(linearity_linear[["sim_1"]][[1]])
saveRDS(linearity_linear,file="./Result/Linearity/linear.RDS")

# plot(linear_term_models[[10]]@frame$uranium,fitted(linear_term_models[[10]]))

linearity_sq <- lapply(`sq_term_models`,function(mod){
    model_linear_term <- lmer(log.radon ~ basement + uranium + (basement | county), data = mod@frame)
    model_sq_term <- lmer(log.radon ~ basement + uranium + I(uranium^2) + (basement | county), data = mod@frame)
    anova_result <- anova(model_linear_term,model_sq_term)
    return(c(stats=anova_result$Chisq[2],p_val=anova_result$`Pr(>Chisq)`[2])) # return test stats and p-value
    })
saveRDS(linearity_sq,file="./Result/Linearity/sq.RDS")
```

### Result

```{r nsim}
nsim <- 5000
```

```{r shapiro-normality-test-results}
normal <- readRDS("~/lme-diag-simulations/simulation/Result/Normality/normal.RDS")
normal_p <- lapply(normal,function(x){return(x$p.value)})
normal_reject <- normal_p[c(which(normal_p < 0.05))]
length(normal_reject)/nsim

t <- readRDS("~/lme-diag-simulations/simulation/Result/Normality/t.RDS")
t_p <- lapply(t,function(x){return(x$p.value)})
t_reject <- t_p[c(which(t_p < 0.05))]
length(t_reject)/nsim

exp <- readRDS("~/lme-diag-simulations/simulation/Result/Normality/exp.RDS")
exp_p <- lapply(exp,function(x){return(x$p.value)})
exp_reject <- exp_p[c(which(exp_p < 0.05))]
length(exp_reject)/nsim
```

```{r bp-test-result}
`var-` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BP/var-.RDS")
`var-_p` <- lapply(`var-`,function(x){return(x['p_val'])})
`bpvar-_reject` <- `var-_p`[c(which(`var-_p` < 0.05))]
length(`bpvar-_reject`)/nsim

`var<` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BP/var<.RDS")
`var<_p` <- lapply(`var<`,function(x){return(x['p_val'])})
`bpvar<_reject` <- `var<_p`[c(which(`var<_p` < 0.05))]
length(`bpvar<_reject`)/nsim

`var<>` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BP/var<>.RDS")
`var<>_p` <- lapply(`var<>`,function(x){return(x['p_val'])})
`bpvar<>_reject` <- `var<>_p`[c(which(`var<>_p` < 0.05))]
length(`bpvar<>_reject`)/nsim

`var>` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BP/var>.RDS")
`var>_p` <- lapply(`var>`,function(x){return(x['p_val'])})
`bpvar>_reject` <- `var>_p`[c(which(`var>_p` < 0.05))]
length(`bpvar>_reject`)/nsim

`var><` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BP/var><.RDS")
`var><_p` <- lapply(`var><`,function(x){return(x['p_val'])})
`bpvar><_reject` <- `var><_p`[c(which(`var><_p` < 0.05))]
length(`bpvar><_reject`)/nsim

```

```{r bf-test-result}
`var-` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF/var-.RDS")
`var-_p` <- lapply(`var-`,function(x){return(x['p_val'])})
`bfvar-_reject` <- `var-_p`[c(which(`var-_p` < 0.05))]
length(`bfvar-_reject`)/nsim

`var<` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF/var<.RDS")
`var<_p` <- lapply(`var<`,function(x){return(x['p_val'])})
`bfvar<_reject` <- `var<_p`[c(which(`var<_p` < 0.05))]
length(`bfvar<_reject`)/nsim

`var<>` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF/var<>.RDS")
`var<>_p` <- lapply(`var<>`,function(x){return(x['p_val'])})
`bfvar<>_reject` <- `var<>_p`[c(which(`var<>_p` < 0.05))]
length(`bfvar<>_reject`)/nsim

`var>` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF/var>.RDS")
`var>_p` <- lapply(`var>`,function(x){return(x['p_val'])})
`bfvar>_reject` <- `var>_p`[c(which(`var>_p` < 0.05))]
length(`bfvar>_reject`)/nsim

`var><` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF/var><.RDS")
`var><_p` <- lapply(`var><`,function(x){return(x['p_val'])})
`bfvar><_reject` <- `var><_p`[c(which(`var><_p` < 0.05))]
length(`bfvar><_reject`)/nsim
```

```{r bf4-test-result}
`var-` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF4/var-.RDS")
`var-_p` <- lapply(`var-`,function(x){return(x['p_val'])})
`bfvar-_reject` <- `var-_p`[c(which(`var-_p` < 0.05))]
length(`bfvar-_reject`)/nsim

`var<` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF4/var<.RDS")
`var<_p` <- lapply(`var<`,function(x){return(x['p_val'])})
`bfvar<_reject` <- `var<_p`[c(which(`var<_p` < 0.05))]
length(`bfvar<_reject`)/nsim

`var<>` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF4/var<>.RDS")
`var<>_p` <- lapply(`var<>`,function(x){return(x['p_val'])})
`bfvar<>_reject` <- `var<>_p`[c(which(`var<>_p` < 0.05))]
length(`bfvar<>_reject`)/nsim

`var>` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF4/var>.RDS")
`var>_p` <- lapply(`var>`,function(x){return(x['p_val'])})
`bfvar>_reject` <- `var>_p`[c(which(`var>_p` < 0.05))]
length(`bfvar>_reject`)/nsim

`var><` <- readRDS("~/lme-diag-simulations/simulation/Result/ConstantVariance/BF4/var><.RDS")
`var><_p` <- lapply(`var><`,function(x){return(x['p_val'])})
`bfvar><_reject` <- `var><_p`[c(which(`var><_p` < 0.05))]
length(`bfvar><_reject`)/nsim
```

```{r linear-test-result}
linear <- readRDS("~/lme-diag-simulations/simulation/Result/Linearity/linear.RDS")
linear_p <- lapply(linear,function(x){return(x['p_val'])})
linear_reject <- linear_p[c(which(linear_p < 0.05))]
length(linear_reject)/nsim

sq <- readRDS("~/lme-diag-simulations/simulation/Result/Linearity/sq.RDS")
sq_p <- lapply(sq,function(x){return(x['p_val'])})
sq_reject <- sq_p[c(which(sq_p < 0.05))]
length(sq_reject)/nsim
```