---
title: "Error Hertero + Nonlinear"
author: "Yicheng Shen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(lmtest)
library(STB)
library(qqplotr)
library(stringr)
library(purrr)
library(plyr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(HLMdiag)
library(nlme)
library(qqplotr)
library(PearsonDS)
library(MASS)
library(gridExtra)
library(doParallel)
library(foreach)
options(warn=-1)
.lmerMod_matrices<- HLMdiag:::.lmerMod_matrices
```


# Analysis of Model Residuals when violating constant variance and linearity assumptions together

3 group settings : Same sized / Balanced / Unbalanced

2 Linaerity: Linear / Sqauared

4 Heterscedasticity factors: Constant Variance / 2 / 4 / 8


# Generate two-level datasets and fit models

```{r}
set.seed(202167)
sa_simdata.df<- generate_data(50,0,0.5)
ba_simdata.df<- generate_data(50,1,0.5)
ub_simdata.df<- generate_data(50,2,0.5)
```


# Simulation functions

```{r simulation-function-heter-nonlinear}
sim.hlm_heterlin <- function(simdata.df, helin.dsn) {

sim_design <- simdata.df # Contain id, grouping and two predictors 

b0 <- 5     # Intercept
b1 <- 1       # X_ij slope 
b2 <- 1       # Z_ij slope 
sd_error <- 1  # # std of the error term, default set as e ~ N(0,1)
sd_int   <- 0.5  # random effect of intercept 
sd_slope <- 0.5  # re of slope

re_df <- data.frame(group = 1:50, 
                    intercept = rnorm(50, mean = 0, sd = sd_int),
                    slope = rnorm(50, mean = 0, sd = sd_slope)) # set up REs for intercept and slope 


    if(helin.dsn == "linear_homo") {   # no change
        sim_df <- left_join(sim_design, re_df, by = "group") %>%
  mutate(error = rnorm(nrow(.), mean = 0, sd = sqrt(sd_error^2 + 0*(X_ij-min(X_ij)))),
         Y_ij = b0 + b1 * X_ij + b2 * Z_ij + intercept + slope * X_ij + error)
    } 

    if(helin.dsn == "sq_2") {  
        sim_df <- left_join(sim_design, re_df, by = "group") %>%
  mutate(error = rnorm(nrow(.), mean = 0, sd = sqrt(sd_error^2 + 2*(X_ij-min(X_ij)))),
         Y_ij = b0 + b1 * X_ij + b2 * (Z_ij)^2 + intercept + slope * X_ij + error)
    } 

    if(helin.dsn == "sq_4") {  
        sim_df <- left_join(sim_design, re_df, by = "group") %>%
  mutate(error = rnorm(nrow(.), mean = 0, sd = sqrt(sd_error^2 + 4*(X_ij-min(X_ij)))),
         Y_ij = b0 + b1 * X_ij + b2 * (Z_ij)^2 + intercept + slope * X_ij + error)
    } 

    if(helin.dsn == "sq_8") {  
        sim_df <- left_join(sim_design, re_df, by = "group") %>%
  mutate(error = rnorm(nrow(.), mean = 0, sd = sqrt(sd_error^2 + 8*(X_ij-min(X_ij)))),
         Y_ij = b0 + b1 * X_ij + b2 * (Z_ij)^2 + intercept + slope * X_ij + error)
    } 
    
    # y.df <- as.data.frame( as.matrix(sim_df$Y_ij) )
    # colnames(y.df) <- paste("sim_", 1:ncol(y.df), sep = "")
    return( sim_df )
}
```


# Hetero + nonlinear Simualtions 

```{r}
helin.dsn <- c("linear_homo","sq_2","sq_4","sq_8") # 1 baseline 3 violations 

set.seed(202167)
nsim <- 20        # default as 20; change as needed

model_resid <- vector("list", length = nsim)
sim_model <- vector("list", length = nsim)
first_resid <- vector("list", length = nsim)
sec_resid <- vector("list", length = nsim)


# same size sim
for(k in seq_along(helin.dsn)){
  
same_size_sim.fm <- replicate (nsim, sim.hlm_heterlin(sa_simdata.df, helin.dsn = helin.dsn[k]))

for(i in seq_along(1:nsim)) {
  sim_model[[i]] <- lmer(Y_ij ~ X_ij + Z_ij + (X_ij | group), data = (same_size_sim.fm)[,i], 
             control = lmerControl(calc.derivs = FALSE))
  
  # level 1 resid
  first_resid[[i]] <- hlm_resid(sim_model[[i]],standardize = T, include.ls = TRUE)
  # level 2 resid
  sec_resid[[i]]<-hlm_resid(sim_model[[i]], level = "group", standardize = T, include.ls = TRUE)
  sec_resid[[i]]$mahalanobis<-mahalanobis_ranef.lmerMod(sim_model[[i]]) 
  # combine into one object
  model_resid[[i]] <-list(first_resid[[i]], sec_resid[[i]])
  names(model_resid[[i]]) = c("first_resid", "second_resid")
}

 saveRDS(model_resid, paste("simulated_models/Result/Heter_nonlinear/", helin.dsn[k], "_same_size_residuals.RDS", sep = ""))
}


# balanced sim
for(k in seq_along(helin.dsn)){

balanced_sim.fm <-  replicate (nsim, sim.hlm_heterlin(ba_simdata.df, helin.dsn = helin.dsn[k]))

for(i in seq_along(1:nsim)) {
  sim_model[[i]] <- lmer(Y_ij ~ X_ij + Z_ij + (X_ij | group), data = balanced_sim.fm[,i], 
             control = lmerControl(calc.derivs = FALSE))

  first_resid[[i]] <- hlm_resid(sim_model[[i]],standardize = T, include.ls = TRUE)

  sec_resid[[i]]<-hlm_resid(sim_model[[i]], level = "group", standardize = T, include.ls = TRUE)
  sec_resid[[i]]$mahalanobis<-mahalanobis_ranef.lmerMod(sim_model[[i]]) 

  model_resid[[i]] <-list(first_resid[[i]], sec_resid[[i]])
  names(model_resid[[i]]) = c("first_resid", "second_resid")
}

 saveRDS(model_resid, paste("simulated_models/Result/Heter_nonlinear/", helin.dsn[k], "_balanced_residuals.RDS", sep = ""))
}


# unbalanced sim
for(k in seq_along(helin.dsn)){
  
unbalanced_sim.fm <-replicate (nsim, sim.hlm_heterlin(ub_simdata.df, helin.dsn = helin.dsn[k]))

for(i in seq_along(1:nsim)) {
  sim_model[[i]] <- lmer(Y_ij ~ X_ij + Z_ij + (X_ij | group), data = unbalanced_sim.fm[,i], 
             control = lmerControl(calc.derivs = FALSE))
  
  first_resid[[i]] <- hlm_resid(sim_model[[i]],standardize = T, include.ls = TRUE)

  sec_resid[[i]]<-hlm_resid(sim_model[[i]], level = "group", standardize = T, include.ls = TRUE)
  sec_resid[[i]]$mahalanobis<-mahalanobis_ranef.lmerMod(sim_model[[i]]) 

  model_resid[[i]] <-list(first_resid[[i]], sec_resid[[i]])
  names(model_resid[[i]]) = c("first_resid", "second_resid")
}

 saveRDS(model_resid, paste("simulated_models/Result/Heter_nonlinear/", helin.dsn[k], "_unbalanced_residuals.RDS", sep = ""))
}

```


# Analysis

```{r load-residuals}
`same_size_linear_var0` <- readRDS("simulated_models/Result/Heter_nonlinear/linear_homo_same_size_residuals.RDS")
`same_size_sq_var2` <- readRDS("simulated_models/Result/Heter_nonlinear/sq_2_same_size_residuals.RDS")
`same_size_sq_var4` <- readRDS("simulated_models/Result/Heter_nonlinear/sq_4_same_size_residuals.RDS")
`same_size_sq_var8` <- readRDS("simulated_models/Result/Heter_nonlinear/sq_8_same_size_residuals.RDS")

`balanced_linear_var0` <- readRDS("simulated_models/Result/Heter_nonlinear/linear_homo_balanced_residuals.RDS")
`balanced_sq_var2` <- readRDS("simulated_models/Result/Heter_nonlinear/sq_2_balanced_residuals.RDS")
`balanced_sq_var4` <- readRDS("simulated_models/Result/Heter_nonlinear/sq_4_balanced_residuals.RDS")
`balanced_sq_var8` <- readRDS("simulated_models/Result/Heter_nonlinear/sq_8_balanced_residuals.RDS")

`unbalanced_linear_var0` <- readRDS("simulated_models/Result/Heter_nonlinear/linear_homo_unbalanced_residuals.RDS")
`unbalanced_sq_var2` <- readRDS("simulated_models/Result/Heter_nonlinear/sq_2_unbalanced_residuals.RDS")
`unbalanced_sq_var4` <- readRDS("simulated_models/Result/Heter_nonlinear/sq_4_unbalanced_residuals.RDS")
`unbalanced_sq_var8` <- readRDS("simulated_models/Result/Heter_nonlinear/sq_8_unbalanced_residuals.RDS")
```


# Linearity

```{r, warning = F}
# Same Sized Grouping anova test
registerDoParallel(cores=2)

linear_var0_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {same_size_linear_var0[[i]]$first_resid$.chol.mar.resid}
sq_var2_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {same_size_sq_var2[[i]]$first_resid$.chol.mar.resid}
sq_var4_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {same_size_sq_var4[[i]]$first_resid$.chol.mar.resid}
sq_var8_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {same_size_sq_var8[[i]]$first_resid$.chol.mar.resid}

linear_var0_fittted <- foreach(i = 1:nsim) %dopar% {same_size_linear_var0[[i]]$first_resid$.mar.fitted}
sq_var2_fittted <-  foreach(i = 1:nsim) %dopar% {same_size_sq_var2[[i]]$first_resid$.mar.fitted}
sq_var4_fittted <-  foreach(i = 1:nsim) %dopar% {same_size_sq_var4[[i]]$first_resid$.mar.fitted}
sq_var8_fittted <-  foreach(i = 1:nsim) %dopar% {same_size_sq_var8[[i]]$first_resid$.mar.fitted}

linear_model_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(linear_var0_chol.mar.resid[[i]] ~ linear_var0_fittted[[i]]),lm(linear_var0_chol.mar.resid[[i]] ~ linear_var0_fittted[[i]] + I(linear_var0_fittted[[i]]^2)))$`Pr(>F)`[2]}

sq_var2_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(sq_var2_chol.mar.resid[[i]] ~ sq_var2_fittted[[i]]),lm(sq_var2_chol.mar.resid[[i]] ~ sq_var2_fittted[[i]] + I(sq_var2_fittted[[i]]^2)))$`Pr(>F)`[2]}

sq_var4_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(sq_var4_chol.mar.resid[[i]] ~ sq_var4_fittted[[i]]),lm(sq_var4_chol.mar.resid[[i]] ~ sq_var4_fittted[[i]] + I(sq_var4_fittted[[i]]^2)))$`Pr(>F)`[2]}

sq_var8_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(sq_var8_chol.mar.resid[[i]] ~ sq_var8_fittted[[i]]),lm(sq_var8_chol.mar.resid[[i]] ~ sq_var8_fittted[[i]] + I(sq_var8_fittted[[i]]^2)))$`Pr(>F)`[2]}

length(linear_model_anova_p[c(which(linear_model_anova_p < 0.05))])/nsim
length(sq_var2_anova_p[c(which(sq_var2_anova_p < 0.05))])/nsim
length(sq_var4_anova_p[c(which(sq_var4_anova_p < 0.05))])/nsim
length(sq_var8_anova_p[c(which(sq_var8_anova_p < 0.05))])/nsim
```

```{r, warning = F}
# Balanced Grouping anova test
registerDoParallel(cores=2)

linear_var0_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {balanced_linear_var0[[i]]$first_resid$.chol.mar.resid}
sq_var2_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {balanced_sq_var2[[i]]$first_resid$.chol.mar.resid}
sq_var4_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {balanced_sq_var4[[i]]$first_resid$.chol.mar.resid}
sq_var8_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {balanced_sq_var8[[i]]$first_resid$.chol.mar.resid}

linear_var0_fittted <- foreach(i = 1:nsim) %dopar% {balanced_linear_var0[[i]]$first_resid$.mar.fitted}
sq_var2_fittted <-  foreach(i = 1:nsim) %dopar% {balanced_sq_var2[[i]]$first_resid$.mar.fitted}
sq_var4_fittted <-  foreach(i = 1:nsim) %dopar% {balanced_sq_var4[[i]]$first_resid$.mar.fitted}
sq_var8_fittted <-  foreach(i = 1:nsim) %dopar% {balanced_sq_var8[[i]]$first_resid$.mar.fitted}

linear_model_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(linear_var0_chol.mar.resid[[i]] ~ linear_var0_fittted[[i]]),lm(linear_var0_chol.mar.resid[[i]] ~ linear_var0_fittted[[i]] + I(linear_var0_fittted[[i]]^2)))$`Pr(>F)`[2]}

sq_var2_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(sq_var2_chol.mar.resid[[i]] ~ sq_var2_fittted[[i]]),lm(sq_var2_chol.mar.resid[[i]] ~ sq_var2_fittted[[i]] + I(sq_var2_fittted[[i]]^2)))$`Pr(>F)`[2]}

sq_var4_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(sq_var4_chol.mar.resid[[i]] ~ sq_var4_fittted[[i]]),lm(sq_var4_chol.mar.resid[[i]] ~ sq_var4_fittted[[i]] + I(sq_var4_fittted[[i]]^2)))$`Pr(>F)`[2]}

sq_var8_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(sq_var8_chol.mar.resid[[i]] ~ sq_var8_fittted[[i]]),lm(sq_var8_chol.mar.resid[[i]] ~ sq_var8_fittted[[i]] + I(sq_var8_fittted[[i]]^2)))$`Pr(>F)`[2]}

length(linear_model_anova_p[c(which(linear_model_anova_p < 0.05))])/nsim
length(sq_var2_anova_p[c(which(sq_var2_anova_p < 0.05))])/nsim
length(sq_var4_anova_p[c(which(sq_var4_anova_p < 0.05))])/nsim
length(sq_var8_anova_p[c(which(sq_var8_anova_p < 0.05))])/nsim
```


```{r, warning = F}
# Unbalanced Grouping anova test
registerDoParallel(cores=2)

linear_var0_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {unbalanced_linear_var0[[i]]$first_resid$.chol.mar.resid}
sq_var2_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {unbalanced_sq_var2[[i]]$first_resid$.chol.mar.resid}
sq_var4_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {unbalanced_sq_var4[[i]]$first_resid$.chol.mar.resid}
sq_var8_chol.mar.resid<- foreach(i = 1:nsim) %dopar% {unbalanced_sq_var8[[i]]$first_resid$.chol.mar.resid}

linear_var0_fittted <- foreach(i = 1:nsim) %dopar% {unbalanced_linear_var0[[i]]$first_resid$.mar.fitted}
sq_var2_fittted <-  foreach(i = 1:nsim) %dopar% {unbalanced_sq_var2[[i]]$first_resid$.mar.fitted}
sq_var4_fittted <-  foreach(i = 1:nsim) %dopar% {unbalanced_sq_var4[[i]]$first_resid$.mar.fitted}
sq_var8_fittted <-  foreach(i = 1:nsim) %dopar% {unbalanced_sq_var8[[i]]$first_resid$.mar.fitted}

linear_model_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(linear_var0_chol.mar.resid[[i]] ~ linear_var0_fittted[[i]]),lm(linear_var0_chol.mar.resid[[i]] ~ linear_var0_fittted[[i]] + I(linear_var0_fittted[[i]]^2)))$`Pr(>F)`[2]}

sq_var2_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(sq_var2_chol.mar.resid[[i]] ~ sq_var2_fittted[[i]]),lm(sq_var2_chol.mar.resid[[i]] ~ sq_var2_fittted[[i]] + I(sq_var2_fittted[[i]]^2)))$`Pr(>F)`[2]}

sq_var4_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(sq_var4_chol.mar.resid[[i]] ~ sq_var4_fittted[[i]]),lm(sq_var4_chol.mar.resid[[i]] ~ sq_var4_fittted[[i]] + I(sq_var4_fittted[[i]]^2)))$`Pr(>F)`[2]}

sq_var8_anova_p <- foreach(i = 1:nsim) %dopar% {anova(lm(sq_var8_chol.mar.resid[[i]] ~ sq_var8_fittted[[i]]),lm(sq_var8_chol.mar.resid[[i]] ~ sq_var8_fittted[[i]] + I(sq_var8_fittted[[i]]^2)))$`Pr(>F)`[2]}

length(linear_model_anova_p[c(which(linear_model_anova_p < 0.05))])/nsim
length(sq_var2_anova_p[c(which(sq_var2_anova_p < 0.05))])/nsim
length(sq_var4_anova_p[c(which(sq_var4_anova_p < 0.05))])/nsim
length(sq_var8_anova_p[c(which(sq_var8_anova_p < 0.05))])/nsim
```


# Constant variance

```{r}
registerDoParallel(cores=2)

# same-sized 
`same_size_tests_linear0` <- foreach(i = 1:nsim) %dopar% {bp((`same_size_linear_var0`[[i]]$first_resid$.std.resid),(`same_size_linear_var0`[[i]]$first_resid$.fitted))}
`var0_p` <- lapply(`same_size_tests_linear0`,function(x){return(x['p_val'])})
`bpvar0_reject` <- `var0_p`[c(which(`var0_p` < 0.05))]
length(`bpvar0_reject`)/nsim

`same_size_tests_sq2` <- foreach(i = 1:nsim) %dopar% {bp((`same_size_sq_var2`[[i]]$first_resid$.std.resid),(`same_size_sq_var2`[[i]]$first_resid$.fitted))}
`var2_p` <- lapply(`same_size_tests_sq2`,function(x){return(x['p_val'])})
`bpvar2_reject` <- `var2_p`[c(which(`var2_p` < 0.05))]
length(`bpvar2_reject`)/nsim

`same_size_tests_sq4` <- foreach(i = 1:nsim) %dopar% {bp((`same_size_sq_var4`[[i]]$first_resid$.std.resid),(`same_size_sq_var4`[[i]]$first_resid$.fitted))}
`var4_p` <- lapply(`same_size_tests_sq4`,function(x){return(x['p_val'])})
`bpvar4_reject` <- `var4_p`[c(which(`var4_p` < 0.05))]
length(`bpvar4_reject`)/nsim

`same_size_tests_sq8` <- foreach(i = 1:nsim) %dopar% {bp((`same_size_sq_var8`[[i]]$first_resid$.std.resid),(`same_size_sq_var8`[[i]]$first_resid$.fitted))}
`var8_p` <- lapply(`same_size_tests_sq8`,function(x){return(x['p_val'])})
`bpvar8_reject` <- `var8_p`[c(which(`var8_p` < 0.05))]
length(`bpvar8_reject`)/nsim
```

```{r}
registerDoParallel(cores=2)

# balanced 
`balanced_var_tests_linear` <- foreach(i = 1:nsim) %dopar% {bp((`balanced_linear_var0`[[i]]$first_resid$.std.resid),(`balanced_linear_var0`[[i]]$first_resid$.fitted))}
`var0_p` <- lapply(`balanced_var_tests_linear`,function(x){return(x['p_val'])})
`bpvar0_reject` <- `var0_p`[c(which(`var0_p` < 0.05))]
length(`bpvar0_reject`)/nsim

`balanced_var_tests_sq2` <- foreach(i = 1:nsim) %dopar% {bp((`balanced_sq_var2`[[i]]$first_resid$.std.resid),(`balanced_sq_var2`[[i]]$first_resid$.fitted))}
`var2_p` <- lapply(`balanced_var_tests_sq2`,function(x){return(x['p_val'])})
`bpvar2_reject` <- `var2_p`[c(which(`var2_p` < 0.05))]
length(`bpvar2_reject`)/nsim

`balanced_var_tests_sq4` <- foreach(i = 1:nsim) %dopar% {bp((`balanced_sq_var4`[[i]]$first_resid$.std.resid),(`balanced_sq_var4`[[i]]$first_resid$.fitted))}
`var4_p` <- lapply(`balanced_var_tests_sq4`,function(x){return(x['p_val'])})
`bpvar4_reject` <- `var4_p`[c(which(`var4_p` < 0.05))]
length(`bpvar4_reject`)/nsim

`balanced_var_tests_sq8` <- foreach(i = 1:nsim) %dopar% {bp((`balanced_sq_var8`[[i]]$first_resid$.std.resid),(`balanced_sq_var8`[[i]]$first_resid$.fitted))}
`var8_p` <- lapply(`balanced_var_tests_sq8`,function(x){return(x['p_val'])})
`bpvar8_reject` <- `var8_p`[c(which(`var8_p` < 0.05))]
length(`bpvar8_reject`)/nsim
```


```{r}
registerDoParallel(cores=2)

# unbalanced 
`unbalanced_var_tests_linear` <- foreach(i = 1:nsim) %dopar% {bp((`unbalanced_linear_var0`[[i]]$first_resid$.std.resid),(`unbalanced_linear_var0`[[i]]$first_resid$.fitted))}
`var0_p` <- lapply(`balanced_var_tests_linear`,function(x){return(x['p_val'])})
`bpvar0_reject` <- `var0_p`[c(which(`var0_p` < 0.05))]
length(`bpvar0_reject`)/nsim

`unbalanced_var_tests_sq2` <- foreach(i = 1:nsim) %dopar% {bp((`unbalanced_sq_var2`[[i]]$first_resid$.std.resid),(`unbalanced_sq_var2`[[i]]$first_resid$.fitted))}
`var2_p` <- lapply(`unbalanced_var_tests_sq2`,function(x){return(x['p_val'])})
`bpvar2_reject` <- `var2_p`[c(which(`var2_p` < 0.05))]
length(`bpvar2_reject`)/nsim

`unbalanced_var_tests_sq4` <- foreach(i = 1:nsim) %dopar% {bp((`unbalanced_sq_var4`[[i]]$first_resid$.std.resid),(`unbalanced_sq_var4`[[i]]$first_resid$.fitted))}
`var4_p` <- lapply(`unbalanced_var_tests_sq4`,function(x){return(x['p_val'])})
`bpvar4_reject` <- `var4_p`[c(which(`var4_p` < 0.05))]
length(`bpvar4_reject`)/nsim

`unbalanced_var_tests_sq8` <- foreach(i = 1:nsim) %dopar% {bp((`unbalanced_sq_var8`[[i]]$first_resid$.std.resid),(`unbalanced_sq_var8`[[i]]$first_resid$.fitted))}
`var8_p` <- lapply(`unbalanced_var_tests_sq8`,function(x){return(x['p_val'])})
`bpvar8_reject` <- `var8_p`[c(which(`var8_p` < 0.05))]
length(`bpvar8_reject`)/nsim
```

