---
output:
  pdf_document: default
  html_document: default
  word_document: default
---
--
title: "Final Project"
output: pdf_document
---

--------------------------------------------


```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, collapse = TRUE, 
                      warning = FALSE, message = FALSE)
library(dplyr)
library(tidyverse)
library(ProbBayes)
library(loo)
library(coda)
library(runjags)
library(MCMCpack)
library(VGAM)
library(bayesrules)
library(Rcpp)
library(runjags)
library(bayesplot)
```


## Data Wrangling

```{r}
#load data
library(readxl)
Covid_Carleton<- read.csv("Covid Data - Carleton.csv")
```


```{r}
#data wrangling -- separate terms
Covid_Carleton <- Covid_Carleton %>% 
  filter(Term == "FA20"| Term == "WI21" | Term == "SP21" | Term == "FA21")  %>%
  mutate(term=recode(Term, "FA20" = 1,"WI21" = 2,"SP21" = 3,"FA21" = 4))
```
## EDA

```{r}
#mutate variable, rename variable
Covid_Summary <- Covid_Carleton %>% 
  mutate(Term = recode(term, "1"="FA20","2"="WI21","3"="SP21","4"="FA21")) %>% 
  group_by(term) %>% 
  summarise(Mean_Positive_Case = mean(Positive_Results),Mean_Positive_Rate_ = mean(Positivity_Rate))

# at most 4 decimal places
knitr::kable(Covid_Summary, digits = 3)
```

```{r}
#standardize variable rice_standard
Covid_Carleton$Rice_Standard <- as.vector(scale(Covid_Carleton$Rice_County))
#EDA 
ggplot(Covid_Carleton, aes(Rice_Standard, Positive_Results, group = 1)) + 
  geom_point() + theme_bw() + 
  geom_point() + geom_smooth(method = "lm", se = F) + 
  xaxis_text(angle = 90, size = 6) +
  ggtitle("Relationship between Carleton Cases and Rice County Cases")

Covid_Carleton$Temp_Standard <- as.vector(scale(Covid_Carleton$Ave_temp))
ggplot(Covid_Carleton, aes(Temp_Standard, Positive_Results, group = 1)) + 
  geom_point() + geom_smooth(method = "lm", se = F) + 
  theme_bw() +
  xaxis_text(angle = 90, size = 6)+
  ggtitle("Relationship between Carleton Cases and Temperature")
```

```{r}
#time series plot of positivity rate in different period
ggplot(Covid_Carleton, aes(Period, Positive_Results, group = 1)) +
  geom_point() +  geom_line() +
  theme_light() +
  xaxis_text(angle = 90, size = 6)+ggtitle("Covid Cases by Week")
```

```{r}
#boxplot of positive rate by term
Covid_Carleton  %>%
  group_by(Term) %>%
  ggplot(aes(Term, Positive_Results)) + geom_boxplot() +
  theme_light() +
  coord_flip()+
  ggtitle("Positive Case by Term")
```


## JAGS

```{r}
# JAGS code
modelString <-"
model{
## sampling
for (i in 1:N){
 y[i] ~ dbin(theta[term[i]], n[i])}
 
## priors
for (j in 1:M){
 theta[j] ~ dbeta(alpha, beta)}
 
alpha <- mu * eta
beta <- (1-mu) *eta
mu ~ dbeta(1,1)
eta <- exp(logeta)
logeta ~ dlogis(log(100), 1)
}"


```


```{r}
#data details
y <- Covid_Carleton$Positive_Results
n <- Covid_Carleton$Total_Tests
N <- length(y)
M <- length(unique(Covid_Carleton$Term))
term <- Covid_Carleton$term
the_data <- list(y = y, n=n, N=N,M=M, term = term) 

# seed for JAGS
init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )

# MCMC draws
carleton_post <- run.jags(
  model = modelString,
  n.chains = 3, 
  data = the_data, 
  monitor = c("theta"), 
  adapt = 1000,  
  burnin = 5000, 
  sample = 10000,
  silent.jags = TRUE  # Eliminates progress bar
)
```

## Diagnostics

```{r}
#convert to mcmc object
post_mcmc <- as.data.frame(as.mcmc(carleton_post))
#trace plot
mcmc_trace(post_mcmc)
#acf plot
mcmc_acf(post_mcmc)
#overlayed density 
mcmc_dens_overlay(carleton_post$mcmc)
#summary mcmc
summary(post_mcmc)
```

## Posterior Predictive
```{r}
# JAGS code

modelString <-"
model{
## sampling
for (i in 1:N){
 y[i] ~ dbin(theta[term[i]], n[i])}
 
## priors
for (j in 1:M){
 theta[j] ~ dbeta(alpha, beta)}
 
alpha <- mu *eta
beta <- (1-mu)*eta
mu ~ dbeta(1,1)
eta <- exp(logeta)
logeta ~ dlogis(log(100), 1)
for (i in 1:N) {
   y_pred[i] ~ dbin(theta[term[i]], n[i])
}

}
"

```


```{r}
# data details
y <- Covid_Carleton$Positive_Results
n <- Covid_Carleton$Total_Tests
N <- length(y)
M <- length(unique(Covid_Carleton$Term))
term <- Covid_Carleton$term
the_data <- list(y = y, n=n, N=N,M=M, term = term) 

#seed 
init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )
#MCMC draws
carleton_pred <- run.jags(
  model = modelString,
  n.chains = 3, 
  data = the_data, 
  monitor = c("y_pred"), 
  adapt = 1000,  
  burnin = 5000, 
  sample = 10000,
  thin=5,
  silent.jags = TRUE  # Eliminates progress bar
)


```

```{r}
#posterior predictive check 
ppc_dens_overlay(y = Covid_Carleton$Positive_Results, yrep = carleton_pred$mcmc[[1]][1:100,])
```

## Inference

```{r}
#make posterior inference plot 
post_intervals <- mcmc_intervals_data(carleton_post$mcmc) %>% 
  mutate(para = c("FA20","WI21","SP21","FA21"))
head(post_intervals)

slice(post_intervals, 1:4) %>%
  ggplot(
    aes(x = reorder(para, (m)), y = (m), ymin = (ll), ymax = (hh))) +
  geom_pointrange() +
  theme_light() +
  xaxis_text(angle = 0, size = 6) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
)+
xlab("Term") + ylab("Positive Test Rate")
```


## Logistical Regression 

```{r}
#JAGS 
modelString <-"
model {
## likelihood
for (i in 1:n){
   y[i] ~ dbin(p[i], N[i])
}
## priors and regression
for (i in 1:n){
   logit(p[i]) <- beta0 +beta1*rice[i]+beta2*temp[i]
   }
## hyperpriors
beta0 ~ dnorm(0, 0.0001)
beta1 ~ dnorm(0, 0.0001)
beta2 ~ dnorm(0, 0.0001)

}
"
# data details
Covid_Carleton <- Covid_Carleton %>% 
  mutate(Rice_Standard = as.vector(scale(Rice_County)),
         temp_standard = as.vector(scale(Ave_temp)))


y <- Covid_Carleton$Positive_Results
n <- length(y)
N <- Covid_Carleton$Total_Tests
rice <- Covid_Carleton$Rice_Standard
temp <- Covid_Carleton$temp_standard
  
the_data <- list(y=y,n=n,N=N,rice=rice,temp=temp)

# seed
init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )

# MCMC draws
carleton_post <- run.jags(
  model = modelString,
  n.chains = 3, 
  data = the_data, 
  monitor = c("beta0","beta1","beta2","y_pred"), 
  adapt = 1000,  
  burnin = 5000, 
  sample = 10000,
  thin=5,
  silent.jags = TRUE  # Eliminates progress bar
)

```

```{r}
# convert to mcmc
post_mcmc <- as.data.frame(as.mcmc(carleton_post)) 
quantile(post_mcmc$beta0,c(0.05,0.95))
quantile(post_mcmc$beta1,c(0.05,0.95))
quantile(post_mcmc$beta2,c(0.05,0.95))
```

```{r}
# diagnostics
#trace plot
mcmc_trace(post_mcmc)
#acf 
mcmc_acf(post_mcmc)
#density plot
mcmc_dens_overlay(carleton_post$mcmc)
#summary
summary(post_mcmc)
```

## Posterior Predictive 

```{r}
#JAGS
modelString <-"
model {
## likelihood
for (i in 1:n){
   y[i] ~ dbin(p[i], N[i])
}
## priors and regression
for (i in 1:n){
   logit(p[i]) <- beta0 +beta1*rice[i]+beta2*temp[i]
   }
## hyperpriors
beta0 ~ dnorm(0, 0.0001)
beta1 ~ dnorm(0, 0.0001)
beta2 ~ dnorm(0, 0.0001)

for (i in 1:n) {
   y_pred[i] ~ dbin(p[i], N[i])
}

}
"
# MCMC draws
carleton_pred <- run.jags(
  model = modelString,
  n.chains = 3, 
  data = the_data, 
  monitor = c("y_pred"), 
  adapt = 1000,  
  burnin = 5000, 
  sample = 10000,
  thin=5,
  silent.jags = TRUE  # Eliminates progress bar
)
```


```{r}
#posterior preditive check 
ppc_dens_overlay(y = Covid_Carleton$Positive_Results, yrep = carleton_pred$mcmc[[1]][1:100,])
```

## Residual Diagnostics 

```{r}
#residual of SLR 
resids <- Covid_Carleton %>%
mutate(
beta0 = mean(post_mcmc$beta0),
beta1 = mean(post_mcmc$beta1),
beta2 = mean(post_mcmc$beta2),
phat_logit = beta0 + beta1 * Rice_Standard + beta2 * temp_standard,
phat = exp(phat_logit)/(1+exp(phat_logit)),
yhat = rbinom(39,Total_Tests, phat),
resid = Positive_Results - yhat
)

# residuals against fitted values
ggplot(data = resids, aes(x = yhat, y = resid)) + # fitted values
  geom_hline(yintercept = 0, linetype = 2, color = "blue") +
  geom_point() + theme_bw()
qqnorm(resids$resid)
qqline(resids$resid)

#against explanatory variable
ggplot(data = resids, aes(x = Rice_Standard, y = resid)) + # fitted values
  geom_hline(yintercept = 0, linetype = 2, color = "blue") +
  geom_point() + theme_bw()
```


```{r}
#quantile of parameter
quantile(post_mcmc$beta2,c(0.05,0.95))
```


# Between instituiton comparison 


```{r model mu, echo=FALSE, out.width='50%'}
# load data
between<- read.csv("proj_11.14.csv")
cross<- between %>% mutate(Test_Rate = Total_Tests/Enrollment)
cross<- cross %>% mutate(type_num = ifelse(Type  == "LAC", 1, 0))
```

## EDA

```{r EDA, echo = F, out.width='50%', fig.align="center"}
library(patchwork)
#plot positive rate vs test rate
a<- ggplot(cross, aes( x = Positive_Rate, y = Test_Rate, group = 1)) +geom_point()+geom_smooth(method = "lm",se = F) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

#plot positive rate vs enrollmeny
b<- ggplot(cross, aes( x = Positive_Rate, y = Enrollment, group = 1)) +geom_point()+geom_smooth(method = "lm",se = F) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 

a+b+plot_layout(ncol = 1, nrow = 2)


```

## model mu (Model2)


```{r}
#JAGS
modelString_mu <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta1*type[i] +beta3*test_rate[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.0001)
beta1 ~ dnorm(0, 0.0001)
beta3 ~ dnorm(0, 0.0001)

## predictive 
for (i in 1:N) {
   y_pred[i] ~ dbin(p[i], n[i])
}

}

"
# seed
init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



# data details
y <- cross$Positive_Results
n <- cross$Total_Tests
type <- cross$type_num
test_rate<- cross$Test_Rate
N <- length(y)
the_data2 <- list("y" = y, "n" = n, "N" = N, "type" = type, "test_rate" = test_rate,
                 "logn" = log(100))


#mcmc draws
posterior_mu <- run.jags(modelString_mu,
                 data = the_data2,
                 n.chains = 3,
                 monitor = c("beta0","beta1","beta3", "p"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 10000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)

posterior_mu_pred <- run.jags(modelString_mu,
                 data = the_data2,
                 n.chains = 3,
                 monitor = c("y_pred"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 10000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)
```

## model p standardized (Model1)

```{r model p stand, results='hide', echo=FALSE}
#JAGS 
modelString_p_stand <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
   ## priors and regression
   logit(p[i]) <- beta0 +beta1*type[i]+beta2*z_enrol[i]+beta3*z_test_rate[i]
}

## hyperpriors

beta0 ~ dnorm(0, 0.0001)
beta1 ~ dnorm(0, 0.0001)
beta2 ~dnorm(0,0.0001)
beta3 ~ dnorm(0, 0.0001)

## predictive 
for (i in 1:N) {
   y_pred[i] ~ dbin(p[i], n[i])
}


}

"

#data details
y <- cross$Positive_Results
n <- cross$Total_Tests
type <- cross$type_num
z_enrol <- (cross$Enrollment -mean(cross$Enrollment))/sqrt(var(cross$Enrollment))
z_test_rate<- (cross$Test_Rate - mean(cross$Test_Rate))/sqrt(var(cross$Test_Rate))
N <- length(y)
the_data4 <- list("y" = y, "n" = n, "N" = N, "type" = type, "z_test_rate" = z_test_rate, "z_enrol" = z_enrol)


#seed 
init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )


#mcmc draws
posterior_p_stand<- run.jags(modelString_p_stand,
                 data = the_data4,
                 n.chains = 3,
                 monitor = c("beta0", "beta1","beta2", "beta3", "p"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 10000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)


posterior_p_stand_pred<- run.jags(modelString_p_stand,
                 data = the_data4,
                 n.chains = 3,
                 monitor = c("y_pred"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 10000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)
```


## model diagnostics

```{r}
# diagnostics for model p standradized
mcmc_trace(posterior_p_stand$mcmc)
mcmc_acf(posterior_p_stand$mcmc)
mcmc_dens_overlay(posterior_p_stand$mcmc)
```

```{r}
# diagnostics for model mu
mcmc_trace(posterior_mu$mcmc)
mcmc_acf(posterior_mu$mcmc)
mcmc_dens_overlay(posterior_mu$mcmc)
```

```{r gelman, geweke, effective sample size, time-series se for model mu}
# gelman -- good
gelman.diag(posterior_mu$mcmc)

# geweke -- good 
geweke.diag(posterior_mu$mcmc)

# effective sample size -- good
effectiveSize(posterior_mu$mcmc)

# pretty good
summary(posterior_mu$mcmc)

```

### Posterior predictive check

```{r}
#not good
ppc_dens_overlay(cross$Positive_Results, posterior_p_stand_pred$mcmc[[1]][1:100, ])
#kinda bad
ppc_stat_grouped(y = cross$Positive_Results, 
                 yrep= as.matrix(posterior_p_stand_pred$mcmc[[1]]),
                 group = cross$Institution)
```

```{r}
ppc_dens_overlay(cross$Positive_Results, posterior_mu_pred$mcmc[[1]][1:100, ])
#actually pretty good
ppc_stat_grouped(y = cross$Positive_Results, 
                 yrep= as.matrix(posterior_mu_pred$mcmc[[1]]),
                 group = cross$Institution)
```



## inference plot

```{r, echo = F, fig.align="center"}
# intervals
post_intervals <- mcmc_intervals_data(posterior_mu$mcmc[, 4:33], regex_pars = "p", prob_outer = 0.9) %>% 
  mutate(para = c("Carleton College","St. Olaf College","University of Minnesota" ,"Macalester College","Smiths College","Harvard University","University of Chicago",
                  "Ohio State University","Johns Hopkins University","University of California LA", "University of California Berkely", "University of Michigan","Boston University",
                  "University of Washington","Rice University","University of Texas Austin","University of Miami","Pennsylvania State University","Amherst College",
                  "Williams College","Middlebury College","Bryn Mawr College","Washington and Lee University",
                  "Bowdoin College","University of Illinois","Purdue University","Colby College","University of Arizona", "Stanford University", "University of Pennsylvania"))

#interval plot
slice(post_intervals, 1:30) %>%
  ggplot(
    aes(x = reorder(para, (m)), y = (m), ymin = (ll), ymax = (hh))) +
  geom_pointrange() +
  theme_light() +
  xaxis_text(angle = 0, size = 6) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
)+ xlab("School") + ylab("Positive Test Rate") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```


### observed rate vs posterior intervals

```{r}
# intervals 
post_intervals_mu <- mcmc_intervals_data(posterior_mu$mcmc, regex_pars = "p", prob_outer = 0.9)

# plot observed rate vs poterior intervals
ggplot(cross, mapping =aes(x = (cross$Positive_Results)/(cross$Total_Tests))) +
  geom_point(data.frame(post_intervals_mu[1:30, ]), 
             mapping = aes(y = post_intervals_mu[1:30,  7] %>% unlist()), alpha = 0.5) +
  geom_linerange(data.frame(post_intervals_mu[1:30, ]), 
                 mapping = aes(ymin = post_intervals_mu[1:30,  5] %>% unlist(), 
                               ymax =post_intervals_mu[1:30,  9] %>% unlist()), alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "dodgerblue")

```



## residual diagnostics for Model1 and Model2

```{r}
# residual diagnostics for model p standardized
post_p_stand<-as.mcmc(posterior_p_stand)
# post_p_stand[, 5:34]
resids_p_stand <- cross %>%
mutate(
   phat = colMeans(post_p_stand[, 5:34]),
   yhat = rbinom(30,cross$Total_Tests, phat),
   resid = Positive_Results - yhat
)

# vs fitted values
ggplot(data = resids_p_stand, aes(x = yhat, y = resid)) + # fitted values
  geom_hline(yintercept = 0, linetype = 2, color = "blue") +
  geom_point() + theme_bw()
qqnorm(resids_p_stand$resid)
qqline(resids_p_stand$resid)

# vs z_test rate
ggplot(data = resids_p_stand, aes(x = z_test_rate, y = resid)) + # fitted values
  geom_hline(yintercept = 0, linetype = 2, color = "blue") +
  geom_point() + theme_bw()


# vs z_enrol
ggplot(data = resids_p_stand, aes(x = z_enrol, y = resid)) + # fitted values
  geom_hline(yintercept = 0, linetype = 2, color = "blue") +
  geom_point() + theme_bw()

## qqplot of model p stand
qqnorm(resids_p_stand$resid)
qqline(resids_p_stand$resid)

```


```{r}
# resiudal diagnostics for model mu
post_mu<-as.mcmc(posterior_mu)
# post_mu[, 4:33]
resids_mu <- cross %>%
mutate(
   phat = colMeans(post_mu[, 4:33]),
   yhat = rbinom(30,cross$Total_Tests, phat),
   resid = Positive_Results - yhat
)

# vs fitted values
ggplot(data = resids_mu, aes(x = yhat, y = resid)) + # fitted values
  geom_hline(yintercept = 0, linetype = 2, color = "blue") +
  geom_point() + theme_bw()
qqnorm(resids_mu$resid)
qqline(resids_mu$resid)

# vs test rate
ggplot(data = resids_mu, aes(x = cross$Test_Rate, y = resid)) + # fitted values
  geom_hline(yintercept = 0, linetype = 2, color = "blue") +
  geom_point() + theme_bw()

```




